Reviving stochasticity: SMBH-binary eccentricity is inherently random

### 2023-06-22 papers/paper-eccentricity/paper-eccentricity.tex
Adjust toy model parameters for better fit

\begin{figure*}
    \centering
    \includegraphics{theta_e_sim_and_model}
    \caption{
    Scattering deflection angle $\theta_\mathrm{defl}$ and the resulting binary eccentricity
    in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
    compared to reasonably well fitting model curves.
    The solid lines show the results from the toy model with $e_\mathrm{s}=0.905$,
    while the semitransparent lines show the results for  $e_\mathrm{s}=0.900$ and $e_\mathrm{s}=0.910$.
    The model curves have been shifted left by $12\degr$ and $6\degr$
    to better match the simulation data, with the shift also indicated by arrows.
    The marginal histograms show kernel density estimates to the simulation data.
    }
    \label{fig:theta_e}
\end{figure*}

@@

The dynamical friction of a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
a Maxwellian distribution with a constant velocity dispersion
${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
set based on the value measured from the simulations \citep[e.g.][]{binney2008}:
%This formula gives the acceleration of a single BH due to dynamical friction as
%\citep[e.g.][]{binney2008}
\begin{equation}\label{eq:chandra_df}
\begin{aligned}
\vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1}.
\end{aligned}
\end{equation}
Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
$X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
which is derived from the size of the simulated systems.

@@

To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
we specify the initial conditions as 
\begin{align}
\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
\dot{\vb{x}}_0 &= (-v_0, 0).
\end{align}
In order to match the typical SMBH relative velocity seen at this separation in
the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ for the $e_0=0.90$ case, and
 $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
%we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
%typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
%and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
due to stellar components that remain tightly bound to the SMBHs and obscure the background
potential relevant for the dynamics,
as well as due to fact that the potential in the simulations is constantly evolving.
However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
eccentricities to be produced,
as is shown by \autoref{fig:model_orbits}.
Performing calculations with different values of $e_\mathrm{s}$,
we found a good fit to the simulation results for $e_\mathrm{s}\approx 0.9$.

@@

The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
These shifts are required due to the rotation of the stellar component tilting the background potential
relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
due to the less radial merger orbit.
In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
although the model does appear to capture the mean behaviour fairly well.
The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
which might be related to the large scatter in the simulation data.

>>>>

\begin{figure*}
    \centering
    \includegraphics{theta_e_sim_and_model}
    \caption{
    Scattering deflection angle $\theta_\mathrm{defl}$ and the resulting binary eccentricity
    in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
    compared to reasonably well fitting model curves.
    The solid lines show the results from the toy model with $e_\mathrm{s}=0.91$,
    while the semitransparent lines show the results for  $e_\mathrm{s}=0.90$ and $e_\mathrm{s}=0.92$.
    The model curves have been shifted left by $12\degr$ and $6\degr$
    to better match the simulation data, with the shift also indicated by arrows.
    The marginal histograms show kernel density estimates to the simulation data.
    }
    \label{fig:theta_e}
\end{figure*}

@@

The dynamical friction of a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
\begin{equation}\label{eq:chandra_df}
\begin{aligned}
\vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
                         &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
\end{aligned}
\end{equation}
where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
and $X = |\vb{v}_1|/\sqrt{2}\sigma_\star$.
We set ${\sigma_\star=200\,\mathrm{km\,s^{-1}}}$ based on the value measured from the simulations.
The value of the Coulomb logarithm is expected to be $\ln{\Lambda}\sim 4\textnormal{--}5$
based on the size of the stellar system,
and we find $\ln{\Lambda} = 4.7$ to give results that agree well with the simulations.

@@

To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
we specify the initial conditions as 
\begin{align}
\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
\dot{\vb{x}}_0 &= (-v_0, 0).
\end{align}
In order to match the typical SMBH relative velocity seen at this separation in
the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ for the $e_0=0.90$ case, and
 $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
due to stellar components that remain tightly bound to the SMBHs and obscure the background
potential relevant for the dynamics,
as well as due to fact that the potential in the simulations is constantly evolving.
However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
eccentricities to be produced,
as is shown by \autoref{fig:model_orbits}.
Performing calculations with different values of $e_\mathrm{s}$,
we found a good fit to the simulation results for $e_\mathrm{s}\approx 0.9$.

@@

The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
These shifts are required due to the rotation of the stellar component tilting the background potential
relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
due to the less radial merger orbit.
In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
although the model does appear to capture the mean behaviour fairly well, apart from the largest values of
$\theta_\mathrm{defl}$.
The behaviour of the model in this region is also quite sensitive to the values of $e_\mathrm{s}$
and $\ln{\Lambda}$,
which might be related to the large scatter in the simulation data.
